---
title: "Warm-Up 02 - Stat 133, Fall 2017"
author: "Vitali Shypko"
date: "9/11/2017"
output: github_document
---

**Loading resources**:
```{r}
# load the objects
load('nba2017-salary-points.RData')
```

```{r}
# list the available objects
ls()
```

All the objects described in the data directory are present (_"experience"_, _"player"_, _"points"_, _"points1"_, _"points2"_, _"points3"_, _"position"_, _"salary"_, _"team"_)

---

**Classes of objects in the file**:
```{r}
class(experience)
class(player)
class(points)
class(points1)
class(points2)
class(points3)
class(position)
class(salary)
class(team)
```

**There are no lists, but there are vectors and a factor**.

---

**Object modes**:
```{r}
mode(experience)
mode(player)
mode(points)
mode(points1)
mode(points2)
mode(points3)
mode(position)
mode(salary)
mode(team)
```

---

**Length of all the objects**:
```{r}
length(experience)
length(player)
length(points)
length(points1)
length(points2)
length(points3)
length(position)
length(salary)
length(team)
```

---

## Exploratory Data Analysis

---

### Quantitative Variable

---

**Summary**:
```{r}
summary(points3)
```

---

**Statistics**:

* _Mean (i.e. average)_:
```{r}
mean(points3)
```

* _Standard deviation_:
```{r}
sd(points3)
```

* _Minimum value_:
```{r}
min(points3)
```

* _Maximum value_:
```{r}
max(points3)
```

* _Median_:
```{r}
median(points3)
```

* _Quartiles_:
```{r}
quantile(points3)
```

**The typical values are ~50**.

---

**The spread in each vector**:
```{r}
range(points3)
```

---

**Distribution**:
```{r}
hist(points3)
```

```{r}
boxplot(points3)
```

---

### Qualitative Variable (_team_)

---

**Is the variable factor**?
```{r}
is.factor(team)
```

---

**Frequency table (counts of each category)**:
```{r}
table(team)
```

---

**Relative frequencies (proportions)**:
```{r}
prop.table(table(team))
```

**Looking at the frequencies and relative frequencies, it looks like the distribution is almost even**.

---

**Frequencies with a barchart**:
```{r}
barplot(table(team))
```

---

### Process report

- What things were hard, even though you saw them in class?

> Describing data and graphs is pretty hard for me as a non-stats major.

- What was easy(-ish) even though we haven't done it in class?

> Markdown syntax and features (getting used to it). Also, using new libraries and functions is pretty straightworward in most cases.

- What type of "errors" you struggled with (if any)?

> "Error: attempt to use zero-length variable name". But it seems to knit regardless of it.

- What are the parts you are not fully understanding?

> I am not fully understanding how exactly we need to describe graphs and our findings.

- What was the most time consuming part?

> The most time consuming part was using R functions that I haven't encountered before (such as density curve).

- Did you collaborate with other students? If so, with who? In what manner?

> I didn't collaborate with other students.

- Was there any frustrating issue? (e.g. RStudio cryptic error, one or more package not playing nice)

> In the very beginning, RStudio couldn't locate the data file. After I restarted it, the error was gone.